---
title: "Facial Processing and Spatial Ability E1"
author: Mona Zhu
date: Updated February 19, 2019
output: 
  html_document:
    fig_width: 8
    fig_height: 5
    toc: true
    toc_float: true
    theme: cerulean
    highlight: tango
    code_folding: hide
---

<style>
pre code, pre, code {
  overflow-y: auto !important;
  max-height: 20em;
}
</style>

```{r, warning=F, message=F}
# libraries
library(plyr)
library(reshape2)
library(stringr)
library(openxlsx)
library(corrgram)
library(tidyverse)
library(corrplot)
library(sjPlot)
library(psych)
library(lavaan)
library(wesanderson)
```

<!-- # Making Wishlist -->

<!-- ```{r Importing Libraries, warning=F} -->
<!-- # convert character class into numeric class -->
<!-- date<-"21Oct2019" -->

<!-- wishlist<-function(mt){ # process mt in order to create wishlist -->
<!--   raw<-read.csv(mt, stringsAsFactors = F) # import wq questionnaire section as data frame -->
<!--   num<-raw %>%  -->
<!--     mutate_if(is.character,as.numeric) -->
<!--   num$score<-rowMeans(num[c(2:23)]) -->
<!--   num$code=ifelse(is.na(num$score), 0, 1) # 1 if we want to use them on wishlist and 0 if we don't -->
<!--   comp<-num[c(1, 25)] -->
<!--   write.table(comp, paste0("Wishlist_", date, ".txt"), sep=" ", row.names=F, col.names=F) # create a txt file with the information required for the wishlist -->
<!-- } -->

<!-- wishlist("MT Processed_21Oct2019.csv") -->
<!-- ``` -->

Reading in data files:

```{r Data Import}
# read the qualtrics data, removing the 2 extra columns at the top that are unnecessary
raw<-read.csv("Examining Aspects of Visuospatial Processing – E1_February 18, 2020_15.41.csv",
              stringsAsFactors = FALSE)[-c(1:2),] 

# only look at participants that completed the task
# & removed our pilot data (i.e., first 3 rows)
# & mental rotation score when items were identical at rotation = 0 degrees > 80% accuracy
# & removed individuals who self-reported not caring
df<-subset(raw, Progress==100 & id!="" & Amnesty.Coded==1)

```

# Route Tracing Task & Scoring

In this task, participants are told to retrace the path they watched on a video by indicating the intersection they passed through.

```{r Route Scoring, warning=F, message=F}
# correct routes:
CorRoute<-c("AH", "HE", "EF", "FA", "Ac", "CK", "KG", "GB", "BY")

# Obtain only the columns with the route drawing + id
route<-cbind(id=df$id, df[ , grepl( "RouteDraw" , names(df))])
route <- mutate_each(route, funs(toupper)) # turn everything into upper cases

# Create a long data frame
route.raw<-melt(route, id=c(1),
              variable.name = "RouteOrder", value.name = "Response")
route.raw$Response<-car::recode(route.raw$Response, "''= NA") # recode empty cells as NAs
route.l<-route.raw[complete.cases(route.raw),] # remove subjects with "NA" responses
route.l$Response<- str_replace(route.l$Response, "EXIT ", "") # replace the word "exit"

ids<-unique(route.l$id) # create variable called "ids" with all the unique SONA ids
route.c<-NULL # empty data frame

# forloop that calculates the percent correct score for the route trace

for (i in ids){
  sub<-subset(route.l, id==i)
  corr<-CorRoute
  for (j in 1:nrow(sub)){
    if (sub$Response[j]==sub$Response[1]) {
      sub$Path[j] <- ifelse(sub$Response[j]=="A", NA, paste0("A", sub$Response[j]))
    } else if (sub$Response[j]!=sub$Response[1]){
      sub$Path[j] <- paste0(sub$Response[j-1], sub$Response[j])
    }
  }
  for(k in 1:nrow(sub)){
    if (is.na(sub$Path[k])) {
      sub$RouteAcc[k]<-NA
    } else if (sub$Path[k] %in% corr) {
      sub$RouteAcc[k]<-1
      corr<-corr[!corr %in% sub$Path[k]]
    } else {
      sub$RouteAcc[k]<-0
    }
  }  
  route.c<-rbind(route.c, sub)  
}


RouteTrace<-ddply(route.c, c("id"), summarise,
                  RoutePercCorr=sum(RouteAcc, na.rm=T)/9)
```

# CFMT & MR Calculation

```{r Scoring Face Task & Mental Rotation}
df$cfmt<-as.numeric(df$SC9)/72 
df$mr<-as.numeric(df$SC10)/56 
```

# Wayfinding Questionnaire

```{r Scoring Wayfinding Qns}
mt<-read.xlsx("F19_MT_and_PS_Oct21.xlsx", sheet = 1, colNames=T)
wq<-cbind(id=mt$id, mt[ , grepl( "WQ_\\d+" , names(mt))])
wq<-wq[complete.cases((wq)),]
wq$id<-as.character(wq$id)

for (i in 1:nrow(wq)){
  wq$var[i]<-sd(wq[i, 2:23])
}

wq_sub<-subset(wq, id %in% ids)


# reverse scoring 8-15
col<-names(wq_sub)[c(9:16)] # reverse code items 8 - 15 according to Claessen et al's paper
wq_sub[,col] = 8 - wq_sub[,col] # reverse code

# navigation & orientation: 1-3, 6, 7, 16, 18-22
wq_sub$nav<-rowMeans(wq_sub[c(2:4, 7:8, 17, 19:23)])

# spatial anxiety: 8-15 reverse coded
wq_sub$anx<-rowMeans(wq_sub[c(9:16)])

# distance estimation: 4,5,17
wq_sub$dist<-rowMeans(wq_sub[c(5,6,18)])

# total score (sum)
wq_sub$WQ_Total<-rowSums(wq_sub[c(2:23)]) 
```

# Pointing Task

```{r Pointing Task}
PT<-df[c(1357:1362,1407)]
#Bows
PT$PointBows<-as.numeric(PT$PointBows)
PT$Bows.Acc <- abs(48-PT$PointBows)

#Chairs
PT$PointChairs <- as.numeric(PT$PointChairs)
PT$Chairs.Acc <-abs(42-PT$PointChairs)

#hat
PT$PointHat <-as.numeric(PT$PointHat)
PT$Hat.Acc <-abs(65-PT$PointHat)

#flower
PT$PointFlower <-as.numeric(PT$PointFlower)
PT$Flower.Acc <-abs(104-PT$PointFlower)

#flamingo
PT$PointFlamingo <-as.numeric(PT$PointFlamingo)
PT$Flamingo.Acc<-abs(117-PT$PointFlamingo)

#Snowflakes
PT$Q74.3 <-as.numeric(PT$Q74.3)
PT$Snowflakes.Acc <-abs(132-PT$Q74.3)


#Average Degrees Off
PT$Pointing.Acc <- rowMeans(PT[,8:13], na.rm = TRUE)

```



# All Scores

```{r Corrgram}
all<-base::merge(wq_sub[,c("id","var", "nav", "anx", "dist", "WQ_Total")], RouteTrace, by="id")
all<-base::merge(all, PT[,c("id", "Pointing.Acc")], by="id")
all<-base::merge(all, df[,c("id", "Gender", "Prev.Exp", "mr", "cfmt")], by="id")
all<-subset(all, var!=0)

# renaming variables
names(all)<-c("id", "WQ_Variance", "WQ_Navigation", "WQ_Anxiety", "WQ_Distance", "WQ_Total", 
              "RouteRetrace", "LandmarkDirection", "Gender", "Familiarity", "MentalRotation",
              "FaceRecognition")
all<-all[c("id","Gender", "Familiarity", "WQ_Navigation", "WQ_Anxiety", "WQ_Distance",
           "WQ_Total", "RouteRetrace", "LandmarkDirection", "MentalRotation", "FaceRecognition")]
all$Gender<-factor(car::recode(all$Gender, "'1'='Women';'2'='Men';'3'='Other'"))
all$Gender<-factor(all$Gender, levels(all$Gender)[c(1,3,2)])
all$Familiarity<-factor(car::recode(all$Familiarity , "'0'='Not Familiar'; '1'='Familiar'; ''='NA'"))
all$Familiarity<-factor(all$Familiarity, levels(all$Familiarity)[c(3,1,2)])


cormat<-cor(all[c(4:ncol(all))])
res1 <- cor.mtest(all[c(4:ncol(all))], conf.level = .95)

corrgram(all[c(4:ncol(all))],
         lower.panel=panel.pts, upper.panel=panel.conf,
         diag.panel=panel.density)
```



```{r Correlation Plot}
corrplot.mixed(cormat, lower="number", lower.col="black", p.mat = res1$p, sig.level = .05)
```




## regression model

We wanted to follow up on the pattern of observations we observed for the cfmt, mental rotation, and landmark

```{r}
mod1<-lm(FaceRecognition~MentalRotation, data=all)
mod2<-lm(FaceRecognition~LandmarkDirection, data=all)
mod3<-lm(FaceRecognition~MentalRotation+LandmarkDirection, data=all)
mod4<-lm(FaceRecognition~MentalRotation*LandmarkDirection, data=all)

tab_model(mod1, mod2, mod3, mod4)
```


```{r}
plot_model(mod4, type="int")
```


Scatter plot of figure above: 
```{r}
ggplot(all, aes(x = FaceRecognition, y = LandmarkDirection, colour=MentalRotation))+
  geom_point()+
  geom_smooth(method="lm")

```

#### Regression Diagnostics

```{r}
par(mfrow = c(2, 2))
plot(mod4)
```

```{r}
# Cook's distance
plot(mod4, 4)
# Residuals vs Leverage
plot(mod4, 5)
```


```{r}
sub<-all[-c(95, 85, 119, 117, 174, 184),]
mod1<-lm(FaceRecognition~MentalRotation, data=sub)
mod2<-lm(FaceRecognition~LandmarkDirection, data=sub)
mod3<-lm(FaceRecognition~MentalRotation+LandmarkDirection, data=sub)
mod4<-lm(FaceRecognition~MentalRotation*LandmarkDirection, data=sub)

tab_model(mod1, mod2, mod3, mod4)
```


```{r}
plot_model(mod4, type="int")
```


# Exploratory Analysis


### Performance Split on Gender

Gender breakdown: 
```{r}
table(all$Gender)

gen<-melt(all, id=c(1:3), variable.name = "Task")
```

The analysis shows that men generally tend to be better than women when it comes to navigation tasks, and women did better on the facial recognition task

#### Self-Reported Navigation Ability

```{r}
graph1<-ddply(gen[grepl("WQ_[NAD]", gen$Task),], .(id, Task, Gender), summarize, 
              mean=mean(value)
              )
graph2<-ddply(graph1, .(Task, Gender), summarize,   
              dep.avg=mean(mean, na.rm=T),
              dep.sd=sd(mean, na.rm=T),
              dep.count=length(mean), 
              dep.ci=dep.sd*qt(0.975,df=dep.count-1)/sqrt(dep.count))
# print(graph2)
ggplot(graph2, aes(x=Task, y=dep.avg, group=Gender, colour=Gender))+
  geom_point(size=3, position = position_dodge(0))+
  geom_errorbar(aes(ymax=dep.avg+dep.ci, 
                    ymin=dep.avg-dep.ci),
                    width=0, position = position_dodge(0))+
  theme(
    panel.background= element_rect(fill=NA), # transparent panel
    plot.background = element_rect(fill=NA, colour=NA), #transparent background
    panel.grid=element_blank(), # remove panel grid
    axis.ticks.x=element_blank(), # remove tick marks on x-axis
    axis.ticks=element_line(colour="gray70"), # change colour of tick marks
    panel.border = element_rect(fill="transparent", colour="gray70"), # change panel border colour
    legend.background = element_rect(fill = "transparent", colour = "transparent"), # change legend background
    axis.text = element_text(color="#432620"),
    legend.key = element_rect(fill = "transparent", colour = "transparent")
  )+
  scale_shape_manual(values=c(16, 1))+
  ggtitle("Self-Reported Navigation Ability")+
  scale_colour_manual(values=wes_palette(n=3, name="GrandBudapest1"))+
  scale_x_discrete("Subfacets of Navigation", labels=c("Navigation", "Anxiety (Reversed)", "Distance"))+
  scale_y_continuous("Average Rating (1-7 Likert Scale)", breaks=seq(0, 10,1), limits=c(0,7))

```


#### Landmark Location Task

```{r}
graph1<-ddply(gen[grepl("Land", gen$Task),], .(id, Task, Gender), summarize, 
              mean=mean(value)
              )
graph2<-ddply(graph1, .(Task, Gender), summarize,   
              dep.avg=mean(mean, na.rm=T),
              dep.sd=sd(mean, na.rm=T),
              dep.count=length(mean), 
              dep.ci=dep.sd*qt(0.975,df=dep.count-1)/sqrt(dep.count))
# print(graph2)
ggplot(graph2, aes(x=Task, y=dep.avg, group=Gender, colour=Gender))+
  geom_point(size=3, position = position_dodge(0.05))+
  geom_errorbar(aes(ymax=dep.avg+dep.ci, 
                    ymin=dep.avg-dep.ci),
                    width=0, position = position_dodge(0.05))+
  theme(
    panel.background= element_rect(fill=NA), # transparent panel
    plot.background = element_rect(fill=NA, colour=NA), #transparent background
    panel.grid=element_blank(), # remove panel grid
    axis.ticks.x=element_blank(), # remove tick marks on x-axis
    axis.ticks=element_line(colour="gray70"), # change colour of tick marks
    panel.border = element_rect(fill="transparent", colour="gray70"), # change panel border colour
    legend.background = element_rect(fill = "transparent", colour = "transparent"), # change legend background
    axis.text = element_text(color="#432620"),
    legend.key = element_rect(fill = "transparent", colour = "transparent")
  )+
  scale_shape_manual(values=c(16, 1))+
  ggtitle("Landmark Location Task (smaller values is better)")+
  scale_colour_manual(values=wes_palette(n=3, name="GrandBudapest1"))+
  scale_x_discrete("")+
  scale_y_continuous("Difference in Degrees", breaks=seq(0, 100,10), limits = c(0,45))

```

#### Route Retrace, Mental Rotation & Face Processing

```{r}
graph1<-ddply(gen[grepl("^[RMF]", gen$Task),], .(id, Task, Gender), summarize, 
              mean=mean(value)
              )
graph2<-ddply(graph1, .(Task, Gender), summarize,   
              dep.avg=mean(mean, na.rm=T),
              dep.sd=sd(mean, na.rm=T),
              dep.count=length(mean), 
              dep.ci=dep.sd*qt(0.975,df=dep.count-1)/sqrt(dep.count))
# print(graph2)
ggplot(graph2, aes(x=Task, y=dep.avg, group=Gender, colour=Gender))+
  geom_point(size=3, position = position_dodge(0.05))+
  geom_errorbar(aes(ymax=dep.avg+dep.ci, 
                    ymin=dep.avg-dep.ci),
                    width=0, position = position_dodge(0.05))+
  theme(
    panel.background= element_rect(fill=NA), # transparent panel
    plot.background = element_rect(fill=NA, colour=NA), #transparent background
    panel.grid=element_blank(), # remove panel grid
    axis.ticks.x=element_blank(), # remove tick marks on x-axis
    axis.ticks=element_line(colour="gray70"), # change colour of tick marks
    panel.border = element_rect(fill="transparent", colour="gray70"), # change panel border colour
    legend.background = element_rect(fill = "transparent", colour = "transparent"), # change legend background
    axis.text = element_text(color="#432620"),
    legend.key = element_rect(fill = "transparent", colour = "transparent")
  )+
  scale_shape_manual(values=c(16, 1))+
  ggtitle("Route Retrace, Mental Rotation & Face Processing Tasks")+
  scale_colour_manual(values=wes_palette(n=3, name="GrandBudapest1"))+
  scale_x_discrete("Tasks")+
  scale_y_continuous("Proportion Correct", breaks=seq(0, 1,.1), limits = c(0,1))

```


Since there aren't enough men in our sample, we next examine whether the pattern of correlation hold for our women sample

```{r}
women<-subset(all, Gender=="Women")

cormat<-cor(women[c(4:ncol(women))])
res1 <- cor.mtest(women[c(4:ncol(women))], conf.level = .95)

corrgram(women[c(4:ncol(women))],
         lower.panel=panel.pts, upper.panel=panel.conf,
         diag.panel=panel.density)
```


```{r}
mod1<-lm(FaceRecognition~MentalRotation, data=women)
mod2<-lm(FaceRecognition~LandmarkDirection, data=women)
mod3<-lm(FaceRecognition~MentalRotation+LandmarkDirection, data=women)
mod4<-lm(FaceRecognition~MentalRotation*LandmarkDirection, data=women)

tab_model(mod1, mod2, mod3, mod4)
```

It seems that the effect is more pronounced in women than for men

```{r}
plot_model(mod4, type="int")
```

Just as a comparison, here is the correlation table for men:

```{r}
men<-subset(all, Gender=="Men")

cormat<-cor(men[c(4:ncol(men))])
res1 <- cor.mtest(men[c(4:ncol(men))], conf.level = .95)

corrgram(men[c(4:ncol(men))],
         lower.panel=panel.pts, upper.panel=panel.conf,
         diag.panel=panel.density)
```

```{r}
mod4<-lm(FaceRecognition~MentalRotation*LandmarkDirection, data=men)

plot_model(mod4, type="int")
```

These results suggest that the effect is more pronounced in women and potentially non-existent in men. 


### Familiarity with building

We wanted to explore whether people who were familiar with the building do better in the pointing and navigation tasks.

A breakdown of individuals who were familiar and not familiar with the building

```{r}
table(all$Familiarity)

```

For the most part, it appears that reported familiarity did not affect individuals' performance on the navigation/landmark location tasks



#### Landmark Location Task

```{r}
graph1<-ddply(gen[grepl("Land", gen$Task),], .(id, Task, Familiarity), summarize, 
              mean=mean(value)
              )
graph2<-ddply(graph1, .(Task, Familiarity), summarize,   
              dep.avg=mean(mean, na.rm=T),
              dep.sd=sd(mean, na.rm=T),
              dep.count=length(mean), 
              dep.ci=dep.sd*qt(0.975,df=dep.count-1)/sqrt(dep.count))
# print(graph2)
ggplot(graph2, aes(x=Task, y=dep.avg, group=Familiarity, colour=Familiarity))+
  geom_point(size=3, position = position_dodge(0.05))+
  geom_errorbar(aes(ymax=dep.avg+dep.ci, 
                    ymin=dep.avg-dep.ci),
                    width=0, position = position_dodge(0.05))+
  theme(
    panel.background= element_rect(fill=NA), # transparent panel
    plot.background = element_rect(fill=NA, colour=NA), #transparent background
    panel.grid=element_blank(), # remove panel grid
    axis.ticks.x=element_blank(), # remove tick marks on x-axis
    axis.ticks=element_line(colour="gray70"), # change colour of tick marks
    panel.border = element_rect(fill="transparent", colour="gray70"), # change panel border colour
    legend.background = element_rect(fill = "transparent", colour = "transparent"), # change legend background
    axis.text = element_text(color="#432620"),
    legend.key = element_rect(fill = "transparent", colour = "transparent")
  )+
  scale_shape_manual(values=c(16, 1))+
  ggtitle("Landmark Location Task (smaller values is better)")+
  scale_colour_manual(values=wes_palette(n=3, name="IsleofDogs1"))+
  scale_x_discrete("")+
  scale_y_continuous("Difference in Degrees", breaks=seq(0, 100,10), limits = c(0,45))

```


#### Route Retrace Task 

```{r}
graph1<-ddply(gen[grepl("^Route", gen$Task),], .(id, Task,  Familiarity), summarize, 
              mean=mean(value)
              )
graph2<-ddply(graph1, .(Task, Familiarity), summarize,   
              dep.avg=mean(mean, na.rm=T),
              dep.sd=sd(mean, na.rm=T),
              dep.count=length(mean), 
              dep.ci=dep.sd*qt(0.975,df=dep.count-1)/sqrt(dep.count))
# print(graph2)
ggplot(graph2, aes(x=Task, y=dep.avg, group=Familiarity, colour=Familiarity))+
  geom_point(size=3, position = position_dodge(0.05))+
  geom_errorbar(aes(ymax=dep.avg+dep.ci, 
                    ymin=dep.avg-dep.ci),
                    width=0, position = position_dodge(0.05))+
  theme(
    panel.background= element_rect(fill=NA), # transparent panel
    plot.background = element_rect(fill=NA, colour=NA), #transparent background
    panel.grid=element_blank(), # remove panel grid
    axis.ticks.x=element_blank(), # remove tick marks on x-axis
    axis.ticks=element_line(colour="gray70"), # change colour of tick marks
    panel.border = element_rect(fill="transparent", colour="gray70"), # change panel border colour
    legend.background = element_rect(fill = "transparent", colour = "transparent"), # change legend background
    axis.text = element_text(color="#432620"),
    legend.key = element_rect(fill = "transparent", colour = "transparent")
  )+
  scale_shape_manual(values=c(16, 1))+
  ggtitle("Route Retrace")+
  scale_colour_manual(values=wes_palette(n=3, name="IsleofDogs1"))+
  scale_x_discrete("Tasks")+
  scale_y_continuous("Proportion Correct", breaks=seq(0, 1,.1), limits = c(0,1))
```


















































# Confirmatory Factor Analysis (Wayfinding Questionnaire)

### Check For Careless Responding
```{r}
library(careless)

careless.string <- cbind(wq$id, longstring(wq, avg = T))
boxplot(careless.string[c(3)], main = "Boxplot of Longest & Average String index")

careless.sub<-subset(careless.string, longstr>8 & avgstr>2.5)

wq2<-subset(wq, !(id %in% careless.sub$`wq$id`))
```

### Model 1: Based on Existing Subscales reported in Claessen et al (2016)

```{r efa on wayfinding data}

way.mod <- '
Navigation =~ 
WQ_1_1 +
WQ_1_2 +
WQ_1_3 +
WQ_1_6 +
WQ_1_7 +
WQ_3_2 +
WQ_3_4 +
WQ_3_5 +
WQ_3_6 +
WQ_3_7 +
WQ_3_8 

Anxiety =~
WQ_1_8 +
WQ_1_9 +
WQ_1_10 +
WQ_1_11 +
WQ_2_1 +
WQ_2_2 +
WQ_2_3 +
WQ_3_1 

Distance =~
WQ_1_4 +
WQ_1_5 +
WQ_3_3 
'

fit<-cfa(way.mod, data = wq2, missing="ML")

summary(fit, fit.measures=T)


```

Looking at the model, it doesn't look like it has great fit (CFI ~ .8 whereas it should be >.95; RMSEA is .12 whereas it shold be < .08)

# Exploratory Factor Anaysis

To examine bottom-up factor loadings

```{r parallel test}
parallel<-fa.parallel(wq2[-c(1, 24:26)], fa="pc")
```

```{r efa}
threefactor<-fa(wq[-c(1,24:26)], nfactors=3, rotate="varimax")

print(threefactor$loadings,cutoff = 0.41)
```

There do seem to be 3 factors in total, but very different than the ones reported in Claessen et al 2006

### Second Confirmatory Factor Analysis 

Based on the factor loadings from the EFA

```{r efa on wayfinding data - new proposed model}

way.mod2 <- '
Navigation =~ 
WQ_1_1 +
WQ_1_2 +
WQ_1_3 +
WQ_1_6 +
WQ_1_7 +
WQ_3_2 +
WQ_3_4 +
WQ_3_5 +
WQ_3_6 +
WQ_3_7 +
WQ_3_8 +
WQ_1_4 +
WQ_1_5 +
WQ_3_3 

Anxiety =~
WQ_1_8 +
WQ_1_9 +
WQ_1_10 +
WQ_1_11 

Lost =~
WQ_2_1 +
WQ_2_2 +
WQ_2_3 +
WQ_3_1 
'

fit2<-cfa(way.mod2, data = wq2, missing="ML")

summary(fit2, fit.measures=T)


```


Model fit has improved (CFI ~ .88; RMSEA ~ .09) but still not awesome.














